Storm surge modelling in the Strait of Georgia

### 2014-10-17 RoughWork/manuscript.tex
started conclusions

\begin{abstract}
The Strait of Georgia is a large, semi-enclosed body of water between Vancouver Island and the mainland of British Columbia connected to the Pacific Ocean via the Strait of Juan de Fuca at the south and Johnstone Strait at the north. During the winter months, coastal communities in the Strait of Georgia are at risk to flooding caused by storm surges, a natural hazard that can occur when a strong storm coincides with high tide. This investigation produces storm surge hindcasts using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea. The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) architecture in a regional configuration. The model is evaluated through comparisons of tidal harmonics, tidal current ellipses, and storm surge water levels with observations. Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly. In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas. Surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca. 
%needs a little more work.
\end{abstract}

@@

At the ocean surface, meteorological conditions and fresh water input from rivers are included. The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2013new}. Forecasts from the period of 2002-2011 are available. Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling.

@@

\section{Discussion}\label{sec:diss}

>>>>

\begin{abstract}
The Strait of Georgia is a large, semi-enclosed body of water between Vancouver Island and the mainland of British Columbia connected to the Pacific Ocean via the Strait of Juan de Fuca at the south and Johnstone Strait at the north. During the winter months, coastal communities in the Strait of Georgia are at risk to flooding caused by storm surges, a natural hazard that can occur when a strong storm coincides with high tide. This investigation produces storm surge hindcasts using a three-dimensional numerical ocean model for the Strait of Georgia and the surrounding bodies of water (Strait of Juan de Fuca, Puget Sound, and Johnstone Strait) collectively known as the Salish Sea. The numerical model employs the Nucleus for European Modelling of the Ocean (NEMO) architecture in a regional configuration. The model is evaluated through comparisons of tidal elevation harmonics and storm surge water levels with observations. Important forcing factors contributing to storm surges have been assessed, finding that the surge entering the domain from the Pacific Ocean contributes most significantly. In addition, local wind patterns cause a slight increase in surge amplitude on the mainland side of the Strait of Georgia when compared to the Vancouver Island coastal areas. Surge amplitudes are found to be greater within the Strait of Georgia when compared to amplitudes in the Strait of Juan de Fuca. 
%needs a little more work.
\end{abstract}

@@

At the ocean surface, meteorological conditions and fresh water input from rivers are included. The ocean surface is forced with momentum and heat fluxes from a 33-km global atmospheric reforecasting model suitable for use in ocean modelling \citep{smith2014new}. Forecasts from the period of 2002-2011 are available. Additionally, the inverse barometric effect of the atmospheric pressure is included, an important consideration in storm surge modelling.

@@

\section{Discussion}\label{sec:diss}
In this article, we have configured a three-dimensional numerical ocean model of the Salish Sea for accurate hindcasts of storm surges in the Strait of Georgia. The model is forced with eight tidal constituents at the western and northern open boundaries, as well as sea surface height anomaly from de-tided observations at Tofino and Port Hardy. Additionaly, wind stress and an inverse barometer effect were forced at the ocean surface using an atmopsheric reforecasting product called CGRF \citep{smith2014new}. Storm surge simulations began with zero initial conditions for the currents and sea surface height and were commenced several days prior to the event of interest to spin up the velocity fields.

Hindcasts of significant events in February 2006, December 2006 and Novemeber 2009 were simulated. In general, the model performs well at reproducing the observed water level heights and surge amplitudes, with Willmott skill scores greater than 0.88 in all cases. We have also assessed the relative importance of different forcing factors contributing to the storm surge. The sea surface height anomaly entering the system from the Pacific Ocean contributes most significantly to the surge amplitude (>50 \%), in agreement with previous studies in this region \citep{murty1995storm}. We have also found that the wind and atmospheric conditions can set up local differences in the surge amplitude, resulting in local variations in the spatial distribution of the surge. Generally, surge amplitudes in the Strait of Georgia are higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions. Results from an additional simulation with no tidal forcing but only surge propagation and atmopsheric forcing found that nonlinear effects due to tide-surge interaction were most important after the San Juan and Gulf Islands.

Several factors considered in other storm surge modelling studies have been neglected. Primarly, effects due to wetting and drying have not been included. This is an important factor to consider in shallow, low lying areas where water levels retreat significantly to reveal land. In this article, the four locations used to assess storm surge accurcay are not in shallow regions and so wetting and drying is not considered to be important. However, several communities in the Strait of Gerogia, such as Boundary Bay and Crescent Beach are in shallow regions and are typically most affected by storm surge during the winter season. As such, efforts to include wetting and drying will be considered in the future. Additionally, sea surface height anomaly from Neah Bay will be used in future real time water leve predictions during the storm surge season.

The Strait of Georgia is a semi-enclosed body of water
